---
format: gfm
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    fig.path: "man/figures/README-"
    out.width: "100%"
---

# h3o

<!-- badges: start -->

[![R-CMD-check](https://github.com/JosiahParry/h3o/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JosiahParry/h3o/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

`{h3o}` is a lightweight R package for interacting with [Uber's H3 Geospatial
Indexing system](https://github.com/uber/h3). The R package uses
[extendr](https://extendr.github.io/) to wrap the eponymous [h3o Rust
crate](https://crates.io/crates/h3o), which offers a pure Rust implementation of
H3, so no linking to Uber’s H3 C library. The package is also intended to work
with the [`{sf}`](https://github.com/r-spatial/sf) package for geometric
operations and as a bonus represents the H3 class as
[`{vctrs}`](https://github.com/r-lib/vctrs), so they work seamlessly within a
tidyverse workflow.

## Installation

You can install the release version of `{h3o}` from CRAN with:

``` r
install.packages("h3o")
```

Or you can install the development version from [GitHub](https://github.com/)
with:

``` r
# install.packages("pak")
pak::pak("JosiahParry/h3o")
```

## Example

H3 vectors can be created from `POINT` geometry columns (`sfc` objects) defined
by sf.

```{r}
#| label: example
#| warning: false
library(h3o)
library(dplyr)
library(sf)
library(tibble)

xy <- data.frame(
  x = runif(100, -5, 10),
  y = runif(100, 40, 50)
)

pnts <- st_as_sf(
  xy,
  coords = c("x", "y"),
  crs = 4326
)

pnts |> mutate(h3 = h3_from_points(geometry, 5))
```

H3 vectors also have an `st_as_sfc()` method which allows conversion of H3 cell
indexes into sf `POLYGON`s.

```{r}
# replace geometry
h3_cells <- pnts |>
  mutate(
    h3 = h3_from_points(geometry, 4),
    geometry = st_as_sfc(h3)
  )

# plot the hexagons
plot(st_geometry(h3_cells))
```

H3 cell centroids can be returned using `h3_to_points()`. If `sf` is avilable,
the results will be returned as an `sfc` (sf column) object. Otherwise it will
return a list of `sfg` (sf geometries).

```{r}
# fetch h3 column
h3s <- h3_cells[["h3"]]

# get there centers
h3_centers <- h3_to_points(h3s)

# plot the hexagons with the centers
plot(st_geometry(h3_cells))
plot(h3_centers, pch = 16, add = TRUE, col = "black")
```

`H3Edge` vectors representing the boundaries of H3 cells can be created with
`h3_edges()`, `h3_shared_edge_pairwise()`, and `h3_shared_edge_sparse()`.

```{r}
cell_edges <- h3_edges(h3s[1:3])
cell_edges
```

We've created a list of each cell's edges. We can flatten them using
`flatten_edges()`.

```{r}
cell_edges <- flatten_edges(cell_edges)
cell_edges
```

These can be cast to sfc objects using `st_as_sfc()`.

```{r}
st_as_sfc(cell_edges)
```

Additionally, you can get the vertexes of H3 cell indexes using
`h3_to_vertexes()` which returns an `sfc_MULTIPOINT`.

```{r}
h3_to_vertexes(h3s)
```

## Bench marks

Since h3o is written in Rust, it is very fast.

### Creating polygons

```{r}
h3_strs <- as.character(h3s)
bench::mark(
  h3o = st_as_sfc(h3s),
  h3jsr = h3jsr::cell_to_polygon(h3_strs)
)
```

### Converting polygons to H3 cells:

```{r}
nc <- st_read(system.file("gpkg/nc.gpkg", package = "sf"), quiet = TRUE) |>
  st_transform(4326) |>
  st_geometry()

bench::mark(
  h3o = sfc_to_cells(nc, 5, "centroid"),
  h3jsr = h3jsr::polygon_to_cells(nc, 5),
  check = FALSE
)
```

### Converting points to cells

```{r}
bench::mark(
  h3o = h3_from_points(pnts$geometry, 3),
  h3jsr = h3jsr::point_to_cell(pnts$geometry, 3),
  check = FALSE
)
```

### Retrieve edges

```{r}
bench::mark(
  h3o = h3_edges(h3s),
  h3jsr = h3jsr::get_udedges(h3_strs),
  check = FALSE
)
```

### Get origins and destinations from edges.

```{r}
# get edges for a single location
eds <- h3_edges(h3s[1])[[1]]
# strings for h3jsr
eds_str <- as.character(eds)

bench::mark(
  h3o = h3_edge_cells(eds),
  h3jsr = h3jsr::get_udends(eds_str),
  check = FALSE
)
```